Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-02-05 paper/manuscript.tex
manuscript updates for theory section

\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
$U$ -- potential energy\\
$u$ -- reduced potential energy \\
$\Delta G$ -- free energy \\
$\Delta f$-- reduced (dimensionless) free energy \\
$K_b$ -- binding or association constant \\
$k_B$-- Boltzmann constant \\
$T$ -- temperature \\
$\beta \equiv (k_B T)^{-1}$ -- inverse thermal energy \\
$K_D$ -- dissociation constant \\
$\lambda$ -- alchemical progress parameter \\
$\Gamma$ -- conformation space accessible by simulations \\
$\vec{q}$ -- single configuration\\
$g$ -- statistical inefficiency
\end{tcolorbox}

\begin{tcolorbox}[title=Common terminology, colback=yellow!10!white]
    \emph{ligand} - small molecule binding to a larger host molecule such as a protein (receptor).\\ 
    \emph{lead optimization} - process of designing a drug candidate from an initial lead compound, usually involving optimizing its affinity, possibly selectivity, and various pharmacokinetic parameters.\\
    \emph{thermodynamic cycle} - sequence of thermodynamic processes (like changes in temperature, volume or pressure) in which the initial and final state are equal.\\
\end{tcolorbox}

@@

In practice, the mutation of $A$ to $B$ in carried out by introducing one or more parameters $\vec{\lambda}$ controlling the potential energy function $U(\vec{q};\vec{\lambda})$ such that the potential of compounds $A$ and $B$ is recovered at two particular values $\vec{\lambda}_A$ and $\vec{\lambda}_B$.
Briefly, this is achieved by simulating a ``chimera'' molecule composed of enough atoms to represent both $A$ and $B$.
A subset of the energetic terms in $U(\vec{q};\vec{\lambda})$ is then modulated by $\vec{\lambda}$ so that at $\vec{\lambda}_A$, the atoms that form molecule $A$ are activated and those belonging exclusively to $B$ are non-interacting ``dummy atoms'', while the opposite occur at $\lambda_B$ (see Section~\ref{sec:relative-fe-protocol} for details).

@@

To solve the problem of having poor overlap between the states of interest, multiple intermediate alchemical states are introduced at values $\vec{\lambda}_A = \vec{\lambda}_0, \vec{\lambda}_1, \cdots, \vec{\lambda}_K = \vec{\lambda}_B$ so that each pair of consecutive states $\vec{\lambda}_k, \vec{\lambda}_{k+1}$ share good overlap.
Each intermediate state models a ligand that is neither $A$ nor $B$ but a mix of the two.
Many estimators (e.g., exponential reweighting (EXP)~\cite{zwanzig1954hightemperature} and Bennet's acceptance ratio (BAR)~\cite{bennett1976efficient,shirts2003equilibriuma}) can then be used to compute the free energy as
\begin{equation}
    \Delta G_{\mathrm{env}} = k_BT \sum_{k=0}^{K-1} \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1})
\end{equation}
from samples collected at all the alchemical states $\{\vec{\lambda}_k \}$, where $\Delta f$ is the \emph{unitless free energy difference}
\begin{equation}
    \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1}) = f(\vec{\lambda}_{k+1}) - f(\vec{\lambda}_k) = - \ln \frac{Z(\vec{\lambda}_{k+1})}{Z(\vec{\lambda}_k)} \, .
\end{equation}
While this strategy usually results in sampling thermodynamic states whose Boltzmann distributions is very similar, thus collecting information that is to some degree redundant, some estimators such as the Multistate Bennett acceptance ratio (MBAR)~\cite{shirts2008statisticallya} can exploit similarities between states to improve the precision of the estimates by using the conformations sampled in all alchemical states $\{\vec{\lambda}_k \}$ to compute the free energy difference $\Delta f(\vec{\lambda}_i, \vec{\lambda}_{j})$ between any pair of states $i,j$  (see Section~\ref{subsec:estimators}).

@@

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations?}
\label{sec:step0}
\todo[inline, color={green!20}]{ASJSM: @everyone please review this section}
When starting an alchemical free energy project, a key first step is to decide whether free energy calculations are really the right tool. Particularly, count the cost of your
project: Can you even hope to tackle the problem with available resources and, if successful, will it
be worth it in terms of human and computational cost?
\subsection{How accurate are alchemical free energy calculations?}
\label{subsec:expectation}
%
Current alchemical free energy calculations involving small molecules seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. A small selection of benchmark datasets and case studies can be found in section~\ref{sec:benchmark} at the end of this document.
However, the domain of applicability is a significant concern~\cite{sherborne2016collaborating, cournia2017relative}, especially for relative calculations, which typically require a high quality and usually experimental bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.
%

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well.
For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, but an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}.
Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}. Here the errors refer to the difference in $\Delta G$ after fitting to the mean experimental binding energy. This prompts us to consider another important aspect. It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. In general, the computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis}. Given the need to understand the performance of the system with alchemical free energy calculations, we recommend that retrospective studies for a particular target and a particular chemical series be performed for each particular application case.
%
\subsection{How reproducible are alchemical free energy calculations?}
\label{subsec:reproducible}
Finite computing resources necessarily limit the generated number of uncorrelated samples of potential energy surfaces, and therefore alchemical free energy calculations only give free energy estimates to within finite precision. An important consideration is how reproducible alchemical free energy calculations are in practice. In simple cases such as absolute hydration free energies of small organic molecules, or relative hydration free energy calculations between structurally similar small organic molecules, it should be possible to obtain highly precise estimates with a given software package (with a sample standard deviation under 0.01 kcal/mol).

@@

Note that these issues concern calculations carried out with a single software package, but simulation package variations can introduce additional issues. Such issues of reproducibility of free energy calculations across different simulation packages have attracted attention recently. Greater variability is expected due to methodological differences such as integrators, thermostats, barostats, treatment of long-range electrostatics, and potentially other factors. For absolute and relative hydration free energies of small organic molecules a variability of ca. 0.2 kcal/mol between popular simulation packages has been reported, even after fixes to the underlying simulation packages.\cite{loeffler2018reproducibility} In the recent SAMPL6 SAMPLing challenge a larger variability of 0.3 to 1.0 kcal/mol was noted in the computed absolute binding free energies of host/guest systems even though similar input and simulation parameters were used \cite{rizzi2019sampl6} and, in many cases, single-point energies were identical or nearly so. Further work is needed to ensure reproducibility of alchemical free energy calculations across implementations to guarantee that force-field development efforts lead to transferable potential energy functions. 
%
\subsection{Is my problem suitable for alchemical free energy calculations?}
\label{subsec:suitability}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting accurate binding free energies for a particular target might be
\emph{more} challenging than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume far more resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required or experiments are otherwise cost-prohibitive.

@@

For example, with more involved restraints, sampling at intermediate $\lambda$ values will usually not need to be as extensive but more computational effort must go to computing the restraining free energy.

@@

    \caption{Alchemical intermediates are created by making the potential energy depend on an additional variable $\lambda$ that interpolates between the chemical endpoints. In (\textbf{A}), at $\lambda=0$ the molecule is a fully interacting phenol and at $\lambda=1$,  a fully interacting benzene. (\textbf{B}) shows an illustration of the probability distribution of the potential energies as the switching function takes values of $\lambda=0$ to $\lambda=1$. Intermediates states are required for a sufficient overlap in potential energies to estimate a free energy difference between $\lambda=0$ and $\lambda=1$.
    Softcore potentials provide one of the most efficient families of intermediate pathways, with a $\lambda$ dependence illustrated in (\textbf{C}) and a potential energy dependence shown in (\textbf{D}). Note how as $\lambda$ approaches 0, the energy smoothly approaches zero at all $r$, a necessary requirement for efficient and stable calculations.  }

@@

The key consideration for choosing alchemical pathways is that the intermediate states that a given pathway produces should sample configurational ensembles that change as slowly as possible as $\lambda$ changes, while still managing to go from the initial state to the final state as $\lambda$ goes from 0 to 1.

Another way of stating this is that intermediate states should sample molecular configurations that are as similar as possible to their neighboring states. The more similar the configurations are between intermediate states, the lower the statistical uncertainty is in the estimate of free energy between intervals. This can be proven directly from the BAR and MBAR formulas~\cite{bennett1976efficient,klimovich2015guidelines}, though the exact same principles apply for TI. For a 'good' path to work, sufficient overlap between potential energies is required. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} and \textbf{B} illustrate this. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} shows in a pictorial way a soft-core potential can be applied across different $\lambda$s. Fig.~\ref{fig:fig_what_is_lambda} \textbf{B} illustrates the potential energy distributions at the different $\lambda$ intermediates, with sufficient overlap between neighboring $\lambda$ states to ensure that reweighting estimators such as MBAR can be used for analysis (see Sec.~\ref{subsec:estimators}). The actual transformation is best handled with soft-core potentials of the form shown in Fig.~\ref{fig:fig_what_is_lambda} \textbf{C} and \textbf{B}, with more details given below.

So how do we adjust the potentials between the two end states based on $\lambda$? The simplest possible alchemical pathway is a \textit{linear} pathway:
\begin{equation}
U(\vec{q},\lambda) = \lambda U_0(\vec{q}) + (1-\lambda)U_1(\vec{q})   
\end{equation}
%
so-called because the dependence on $\lambda$ is linear. This clearly satisfies the basic requirement that it gives the initial endpoint potential energy $U_0(\vec{q})$ when $\lambda=0$ and final endpoint energy $U_1(\vec{q})$ when $\lambda=1$.

@@

However, this approach ends up being terrible for removing or adding repulsive potentials that go to infinity quickly at or near the origin. One way to look at this is to examine how low $\lambda$ values must go to reduce the energy at $0.5\sigma$ (the atomic size parameter) down to 1 $k_BT$, where thermal fluctuations make it possible for other atomic sites to penetrate routinely that deep. If we are trying to go from a particle being present, and desire to make it disappear alchemically, then if the repulsive terms are of the form $\epsilon\frac{\sigma}{r}^{12}$, then if $\epsilon$ was 1 $k_BT$ at the temperature of interest, and we start with the particle present, then solving for $(1-\lambda)(1 kB_T)\frac{\sigma}{\frac{1}{2}\sigma}) = 1 k_B T$ we get $\lambda = 1-2^{-12} \sim 0.999976$. So we have gone virtually all the way to the end of the transformation, but there is still an impenetrable post in the middle of our simulation! This is not very much like the desired final state of no interactions between the particle and its environment. We can play around with a few ways of modifying this, like simulating many more intermediate states near $\lambda=1$. However, various analyses have shown that this is not a very good strategy~\cite{pham2011identifying, beutler1994avoiding, zacharias1994separationshifted, blondel2004ensemble, gapsys2012new}.

What we need instead is a function that smoothly gets rid of this infinity. A large number of schemes have been tried~\cite{beutler1994avoiding, zacharias1994separationshifted, blondel2004ensemble, pham2011identifying, pham2012optimal, naden2014linear, donnini2005molecular}, but the most common strategy that appears to be the best practice is to use a "soft-core" potential, of the form:
%
\begin{equation}
    U(\vec{r_{ij}},\lambda) = 4\epsilon_{ij} \lambda \left(\frac{1}{(\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6)^2} -  \frac{1}{\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6}\right)
    \label{eq:softcore}
\end{equation}
%
Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\lambda=1$, and the other endpoint ($\lambda=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\lambda$ goes to zero, the  $\alpha(1-\lambda)$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one  given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in \textbf{C} and \textbf{D} of Fig.~\ref{fig:fig_what_is_lambda}.

@@

We have discussed optimal ways of disappearing or appearing Lennard-Jones interaction sites and turning on and off electrostatics terms. What about performing both transformations at the same time? We can not turn off the electrostatics linearly at the same time we turn off the Lennard-Jones terms, as it would leave infinitely large attractive and repulsive electrostatic terms "bare" at small $\lambda$, resulting in the simulation crashing. It \textit{is} possible to apply the same soft core approach to the Coulomb interaction and this is indeed done in a number of implementations, in which case it is important that the Coulomb interaction is softened more rapidly than the Lennard-Jones interaction to avoid charge penetration issues, this can be tricky to ensure for all types of perturbations ~\cite{steinbrecher2011softcore}.

@@

Convergence in other alchemical observables should also be monitored to determine if the defined phase space has been sufficiently sampled and enough decorrelated samples have been drawn. These additional observables include, but are not limited to, the variance in $\frac{dU}{d\lambda}$ across all $\lambda$ values, calculating the variance in free energy using bootstrap analysis, and comparing differences in free energies calculated using different percentages of the simulation in both the forward and reverse directions ( see Fig. ~\ref{fig:convergence_forward_reverse}).

@@

Each method requires different data from the simulation be collected.  If, for instance, the free energy estimator selected is thermodynamic integration, then values of $\frac{dU}{d\lambda}$ at uncorrelated data points must be collected. Once a combination of knowing what type of simulation you will run, which alchemical topology you will simulate, what alchemical path you will simulation along, and what your stopping conditions are, then you are ready to enumerate the information you should capture. Below is a sample of the minimal information you need for a set of common estimators (discussed in more detail in Sec.~\ref{subsec:estimators}):

@@

    \item Thermodynamic Integration (TI) requires $\frac{\partial U(\vec{x})}{\partial\lambda}$.

@@

If the alchemical path you have chosen is a linear alchemical path, then you get $\frac{dU}{d\lambda} = U_0(\vec{x}) - U_1(\vec{x})$, which is the difference between the initial and final states. 

@@

For simple independent simulations that sample $x_t \sim \pi(x ; \lambda)$, this is given by the reduced potential

@@

a_t &\equiv& - \ln \pi(x_t; \lambda) + c = u(x_t; \lambda) .

@@

If independent simulations are used, the alchemical state $\lambda$ may have a significant impact on the correlation time, and these simulations should be subsampled independently using a separate estimate of the statistical inefficinecy $g$ for each alchemical state.

@@

variable $\lambda$, we find that:

@@

\frac{df}{d\lambda} = \frac{d}{d\lambda} \left[-\ln \int \frac{\exp{-u(\lambda,\vec{q})}}{Q(\lambda)} d\vec{q}\right] = \expect{\frac{du(\lambda,\vec{q})}{d\lambda}\
}_{\lambda} 

@@

And than we can then numerically integrate $df/d\lambda$ over an alchemical transformation, using a range of different well-established techniques, to obtain:

@@

\Delta f    = \int_{0}^{1} \expect{\frac{du(\lambda,\vec{q})}{d\lambda}}_{\lambda}  d\lambda    

@@

This approach to calculating the free energy is called thermodynamic integration (TI). Averaging over $\expect{\frac{du}{d\lambda}}$ requires fewer uncorrelated samples to reach a given level of relative error

>>>>

\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
$L$ and $R$ -- generic names for ligand and receptor\\
$K_b^{\circ}$ -- binding  constant \\
$c^{\circ}$ -- standard state concentration \\
$U$ -- potential energy\\
$u$ -- reduced potential describing a thermodynamic state \\
$\Delta G$ -- Gibbs free energy (isothermal isobaric ensemble)\\
$\Delta A$ -- Helmholtz free energy (canonical ensemble)\\
$\Delta f$-- reduced (dimensionless) free energy \\
$\Delta \hat{f}$ -- estimate from an estimator for the reduced free energy\\
$\Gamma$ -- conformation space accessible by simulations \\
$\vec{q}$ -- vector of a single configuration, i.e. $x$, $y$, $z$ coordinates of the simulation system\\
$k_B$ -- Boltzmann constant \\
$Z$ -- partition function \\
$p$ -- pressure \\
$\mu$ -- chemical potential (grand canonical ensemble)\\
$T$ -- temperature \\
$\beta \equiv (k_B T)^{-1}$ -- inverse thermal energy \\
$\vec{\lambda}$ -- alchemical progress parameter \\
$g$ -- statistical inefficiency
\end{tcolorbox}

\begin{tcolorbox}[title=Common terminology, colback=yellow!10!white]
    \emph{ligand} - small molecule binding to a larger host molecule such as a protein (receptor).\\ 
    \emph{lead optimization} - process of designing a drug candidate from an initial lead compound, usually involving optimizing its affinity, possibly selectivity, and various pharmacokinetic parameters.\\
    \emph{docking} -- Method of generate simulation starting structures for ligand receptor systems, when no crystal structure is available\\
    \emph{thermodynamic cycle} - sequence of thermodynamic processes (like changes in temperature, volume or pressure) in which the initial and final state are equal.\\
    \emph{alchemical} -- states or simulations that contain non-physical intermediates \\
\end{tcolorbox}

@@

In practice, the mutation of $A$ to $B$ in carried out by introducing one or more parameters $\vec{\lambda}$ controlling the potential energy function $U(\vec{q};\vec{\lambda})$ such that the potential of compounds $A$ and $B$ is recovered at two particular values $\vec{\lambda}_A$ and $\vec{\lambda}_B$.
Briefly, this is achieved by simulating a ``chimera'' molecule composed of enough atoms to represent both $A$ and $B$.
A subset of the energetic terms in $U(\vec{q};\vec{\lambda})$ is then modulated by $\vec{\lambda}$ so that at $\vec{\lambda}_A$, the atoms that form molecule $A$ are activated and those belonging exclusively to $B$ are non-interacting ``dummy atoms'', while the opposite occur at $\vec{\lambda}_B$ (see Section~\ref{sec:relative-fe-protocol} for details).

@@

To solve the problem of having poor overlap between the states of interest, multiple intermediate alchemical states are introduced at values $\vec{\lambda}_A = \vec{\lambda}_0, \vec{\lambda}_1, \cdots, \vec{\lambda}_K = \vec{\lambda}_B$ so that each pair of consecutive states $\vec{\lambda}_k, \vec{\lambda}_{k+1}$ share good overlap.
Each intermediate state models a ligand that is neither $A$ nor $B$ but a mix of the two.
Many estimators (e.g., exponential reweighting (EXP)~\cite{zwanzig1954hightemperature} and Bennet's acceptance ratio (BAR)~\cite{bennett1976efficient,shirts2003equilibriuma}) can then be used to compute the free energy as
\begin{equation}
    \Delta G_{\mathrm{env}} = k_BT \sum_{k=0}^{K-1} \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1})
\end{equation}
from samples collected at all the alchemical states $\{\vec{\lambda}_k \}$, where $\Delta f$ is the \emph{unitless free energy difference}
\begin{equation}
    \Delta f(\vec{\lambda}_k, \vec{\lambda}_{k+1}) = f(\vec{\lambda}_{k+1}) - f(\vec{\lambda}_k) = - \ln \frac{Z(\vec{\lambda}_{k+1})}{Z(\vec{\lambda}_k)} \, .
\end{equation}
While this strategy usually results in sampling thermodynamic states whose Boltzmann distributions is very similar, thus collecting information that is to some degree redundant, some estimators such as the Multistate Bennett acceptance ratio (MBAR)~\cite{shirts2008statisticallya} can exploit similarities between states to improve the precision of the estimates by using the conformations sampled at all alchemical states $\{\vec{\lambda}_k \}$ to compute the free energy difference $\Delta f(\vec{\lambda}_i, \vec{\lambda}_{j})$ between any pair of states $i,j$  (see Section~\ref{subsec:estimators}).

@@

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations?}
\label{sec:step0}
When starting an alchemical free energy project, a key first step is to decide whether free energy calculations are really the right tool. Particularly, count the cost of your
project: Can you even hope to tackle the problem with available resources and, if successful, will it
be worth it in terms of human and computational cost?

\subsection{How accurate are alchemical free energy calculations?}
\label{subsec:expectation}
Current alchemical free energy calculations involving small molecules seem to achieve, in favorable cases, root mean square (RMS) errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. A small selection of benchmark datasets and case studies can be found in section~\ref{sec:benchmark} at the end of this document.
However, the domain of applicability is a significant concern~\cite{sherborne2016collaborating, cournia2017relative}, especially for relative calculations, which typically require a high quality and usually experimental bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well.
For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, but an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}.
Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}. Here the errors refer to the difference in $\Delta G$ after fitting to the mean experimental binding energy. This prompts us to consider another important aspect. It is important to be clear on what error to report: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. (See also the section recommendations for best practices of reporting data Sec.~\ref{sec:plot_data}) In general, the computed edge $\Delta \Delta G$ RMSEs significantly underestimates the error of all edge RMSEs: \url{https://github.com/jchodera/jacs-dataset-analysis}. Given the need to understand the performance of the system with alchemical free energy calculations, we recommend that retrospective studies for a particular target and a particular chemical series be performed for each particular application case.

\subsection{How reproducible are alchemical free energy calculations?}
\label{subsec:reproducible}
Finite computing resources necessarily limit the generated number of uncorrelated samples of potential energy surfaces, and therefore alchemical free energy calculations only give free energy estimates to within finite precision. An important consideration is how reproducible alchemical free energy calculations are in practice. In simple cases such as absolute hydration free energies of small organic molecules, or relative hydration free energy calculations between structurally similar small organic molecules, it should be possible to obtain highly precise estimates with a given software package (with a sample standard deviation under 0.01 kcal/mol). [DO WE HAVE A CIATION FOR THIS?]

@@

Note that these issues concern calculations carried out with a single software package, but simulation package variations can introduce additional issues. Such issues of reproducibility of free energy calculations across different simulation packages have attracted attention recently. Greater variability is expected due to methodological differences such as integrators, thermostats, barostats, treatment of long-range electrostatics, and potentially other factors. For absolute and relative hydration free energies of small organic molecules a variability of ca. 0.2 kcal/mol between popular simulation packages has been reported, even after fixes to the underlying simulation packages~\cite{loeffler2018reproducibility}. In the recent SAMPL6 SAMPLing challenge a larger variability of 0.3 to 1.0 kcal/mol was noted in the computed absolute binding free energies of host/guest systems even though similar input and simulation parameters were used~\cite{rizzi2019sampl6} and, in many cases, single-point energies were identical or nearly so. Further work is needed to ensure reproducibility of alchemical free energy calculations across implementations to guarantee that force-field development efforts lead to transferable potential energy functions.

\subsection{Is my problem suitable for alchemical free energy calculations?}
\label{subsec:suitability}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting accurate binding free energies for a particular target might be
\emph{more} challenging than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume far more resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required or experiments are otherwise cost-prohibitive.

@@

For example, with more involved restraints, sampling at intermediate $\vec{\lambda}$ values will usually not need to be as extensive but more computational effort must go to computing the restraining free energy.

@@

    \caption{Alchemical intermediates are created by making the potential energy depend on an additional variable $\vec{\lambda}$ that interpolates between the chemical endpoints. In (\textbf{A}), at $\vec{\lambda}=0$ the molecule is a fully interacting phenol and at $\vec{\lambda}=1$,  a fully interacting benzene. (\textbf{B}) shows an illustration of the probability distribution of the potential energies as the switching function takes values of $\vec{\lambda}=0$ to $\vec{\lambda}=1$. Intermediates states are required for a sufficient overlap in potential energies to estimate a free energy difference between $\vec{\lambda}=0$ and $\vec{\lambda}=1$.
    Softcore potentials provide one of the most efficient families of intermediate pathways, with a $\vec{\lambda}$ dependence illustrated in (\textbf{C}) and a potential energy dependence shown in (\textbf{D}). Note how as $\vec{\lambda}$ approaches 0, the energy smoothly approaches zero at all $r$, a necessary requirement for efficient and stable calculations.  }

@@

The key consideration for choosing alchemical pathways is that the intermediate states that a given pathway produces should sample configurational ensembles that change as slowly as possible as $\vec{\lambda}$ changes, while still managing to go from the initial state to the final state as $\vec{\lambda}$ goes from 0 to 1.

Another way of stating this is that intermediate states should sample molecular configurations that are as similar as possible to their neighboring states. The more similar the configurations are between intermediate states, the lower the statistical uncertainty is in the estimate of free energy between intervals. This can be proven directly from the BAR and MBAR formulas~\cite{bennett1976efficient,klimovich2015guidelines}, though the exact same principles apply for TI. For a 'good' path to work, sufficient overlap between potential energies is required. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} and \textbf{B} illustrate this. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} shows in a pictorial way a soft-core potential can be applied across different $\vec{\lambda}$s. Fig.~\ref{fig:fig_what_is_lambda} \textbf{B} illustrates the potential energy distributions at the different $\vec{\lambda}$ intermediates, with sufficient overlap between neighboring $\vec{\lambda}$ states to ensure that reweighting estimators such as MBAR can be used for analysis (see Sec.~\ref{subsec:estimators}). The actual transformation is best handled with soft-core potentials of the form shown in Fig.~\ref{fig:fig_what_is_lambda} \textbf{C} and \textbf{B}, with more details given below.

So how do we adjust the potentials between the two end states based on $\vec{\lambda}$? The simplest possible alchemical pathway is a \textit{linear} pathway:
\begin{equation}
U(\vec{q},\vec{\lambda}) = \vec{\lambda} U_0(\vec{q}) + (1-\vec{\lambda})U_1(\vec{q})   
\end{equation}
%
so-called because the dependence on $\vec{\lambda}$ is linear. This clearly satisfies the basic requirement that it gives the initial endpoint potential energy $U_0(\vec{q})$ when $\vec{\lambda}=0$ and final endpoint energy $U_1(\vec{q})$ when $\vec{\lambda}=1$.

@@

However, this approach ends up being terrible for removing or adding repulsive potentials that go to infinity quickly at or near the origin. One way to look at this is to examine how low $\vec{\lambda}$ values must go to reduce the energy at $0.5\sigma$ (the atomic size parameter) down to 1 $k_BT$, where thermal fluctuations make it possible for other atomic sites to penetrate routinely that deep. If we are trying to go from a particle being present, and desire to make it disappear alchemically, then if the repulsive terms are of the form $\epsilon\frac{\sigma}{r}^{12}$, then if $\epsilon$ was 1 $k_BT$ at the temperature of interest, and we start with the particle present, then solving for $(1-\vec{\lambda})(1 kB_T)\frac{\sigma}{\frac{1}{2}\sigma}) = 1 k_B T$ we get $\vec{\lambda} = 1-2^{-12} \sim 0.999976$. So we have gone virtually all the way to the end of the transformation, but there is still an impenetrable post in the middle of our simulation! This is not very much like the desired final state of no interactions between the particle and its environment. We can play around with a few ways of modifying this, like simulating many more intermediate states near $\vec{\lambda}=1$. However, various analyses have shown that this is not a very good strategy~\cite{pham2011identifying, beutler1994avoiding, zacharias1994separationshifted, blondel2004ensemble, gapsys2012new}.

What we need instead is a function that smoothly gets rid of this infinity. A large number of schemes have been tried~\cite{beutler1994avoiding, zacharias1994separationshifted, blondel2004ensemble, pham2011identifying, pham2012optimal, naden2014linear, donnini2005molecular}, but the most common strategy that appears to be the best practice is to use a "soft-core" potential, of the form:
%
\begin{equation}
    U(\vec{r_{ij}},\vec{\lambda}) = 4\epsilon_{ij} \vec{\lambda} \left(\frac{1}{(\alpha(1-\vec{\lambda}) + (r_{ij}/\sigma_{ij})^6)^2} -  \frac{1}{\alpha(1-\vec{\lambda}) + (r_{ij}/\sigma_{ij})^6}\right)
    \label{eq:softcore}
\end{equation}
%
Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\vec{\lambda}=1$, and the other endpoint ($\vec{\lambda}=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\vec{\lambda}$ goes to zero, the  $\alpha(1-\vec{\lambda})$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one  given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in \textbf{C} and \textbf{D} of Fig.~\ref{fig:fig_what_is_lambda}.

@@

We have discussed optimal ways of disappearing or appearing Lennard-Jones interaction sites and turning on and off electrostatics terms. What about performing both transformations at the same time? We can not turn off the electrostatics linearly at the same time we turn off the Lennard-Jones terms, as it would leave infinitely large attractive and repulsive electrostatic terms "bare" at small $\vec{\lambda}$, resulting in the simulation crashing. It \textit{is} possible to apply the same soft core approach to the Coulomb interaction and this is indeed done in a number of implementations, in which case it is important that the Coulomb interaction is softened more rapidly than the Lennard-Jones interaction to avoid charge penetration issues, this can be tricky to ensure for all types of perturbations ~\cite{steinbrecher2011softcore}.

@@

Convergence in other alchemical observables should also be monitored to determine if the defined phase space has been sufficiently sampled and enough decorrelated samples have been drawn. These additional observables include, but are not limited to, the variance in $\frac{dU}{d\vec{\lambda}}$ across all $\vec{\lambda}$ values, calculating the variance in free energy using bootstrap analysis, and comparing differences in free energies calculated using different percentages of the simulation in both the forward and reverse directions ( see Fig. ~\ref{fig:convergence_forward_reverse}).

@@

Each method requires different data from the simulation be collected.  If, for instance, the free energy estimator selected is thermodynamic integration, then values of $\frac{dU}{d\vec{\lambda}}$ at uncorrelated data points must be collected. Once a combination of knowing what type of simulation you will run, which alchemical topology you will simulate, what alchemical path you will simulation along, and what your stopping conditions are, then you are ready to enumerate the information you should capture. Below is a sample of the minimal information you need for a set of common estimators (discussed in more detail in Sec.~\ref{subsec:estimators}):

@@

    \item Thermodynamic Integration (TI) requires $\frac{\partial U(\vec{x})}{\partial\vec{\lambda}}$.

@@

If the alchemical path you have chosen is a linear alchemical path, then you get $\frac{dU}{d\vec{\lambda}} = U_0(\vec{x}) - U_1(\vec{x})$, which is the difference between the initial and final states. 

@@

For simple independent simulations that sample $x_t \sim \pi(x ; \vec{\lambda})$, this is given by the reduced potential

@@

a_t &\equiv& - \ln \pi(x_t; \vec{\lambda}) + c = u(x_t; \vec{\lambda}) .

@@

If independent simulations are used, the alchemical state $\vec{\lambda}$ may have a significant impact on the correlation time, and these simulations should be subsampled independently using a separate estimate of the statistical inefficinecy $g$ for each alchemical state.

@@

variable $\vec{\lambda}$, we find that:

@@

\frac{df}{d\vec{\lambda}} = \frac{d}{d\vec{\lambda}} \left[-\ln \int \frac{\exp{-u(\vec{\lambda},\vec{q})}}{Q(\vec{\lambda})} d\vec{q}\right] = \expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}\
}_{\vec{\lambda}} 

@@

And than we can then numerically integrate $df/d\vec{\lambda}$ over an alchemical transformation, using a range of different well-established techniques, to obtain:

@@

\Delta f    = \int_{0}^{1} \expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}}_{\vec{\lambda}}  d\vec{\lambda}    

@@

This approach to calculating the free energy is called thermodynamic integration (TI). Averaging over $\expect{\frac{du}{d\vec{\lambda}}}$ requires fewer uncorrelated samples to reach a given level of relative error
